---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# houser

The US Census Bureau conducts [quite a few](https://www.census.gov/econ/construction.html) 
surveys of buildings and housing in the US. None of this data is available in
the [official Census API](https://www.census.gov/data/developers/data-sets.html)
This package provides a convenient way to access housing data from the US
Census.

## Warning

Right now, `houser` can only download and clean **annual** files from
the Building Permits Survey. **Do not attempt to download other BPS files using `houser`.** 
I have no idea what could happen. If you don't know what annual files are, see
the "Data" section below before using `houser`.

`get_bps()` will download annual files by default, so just don't change the
`formats` argument in `get_bps()` and you'll be fine. `clean_bps()` currently
only cleans annual files.

## Data

Right now, `houser` can only download and clean **annual** (a) files from the
[Building Permits Survey](https://www.census.gov/construction/bps/), in [this directory](https://www2.census.gov/econ/bps/). See the [BPS documentation](https://www2.census.gov/econ/bps/Documentation/) for all the info
you'll ever need. The following is an attempt at a summary.

This table gives you a lay of the land:

|  Geography|                            Formats|         Regions|     Years|
|----------:|----------------------------------:|---------------:|---------:|
|      State|                            c, y, a|             N/A| 1980-2017|
|     County|                            c, y, a|             N/A| 1980-2017|
|        MSA|                            c, y, a|             N/A| 1980-2017|
|      Place|  monthly: c, y, r;<br>annual: a, r|  mw, ne, so, we| 1980-2017|

BPS data are just a bunch of csv files, available for 4 geographies. Geography
is the unit of observation in all BPS files. For each geography, files come in 5
possible formats for every year since 1980 (see "Formats" below), though not all
formats are available for all geographies. Place (municipality)-level files are
organized by the region of the US in which their places are located (see
"Regions" below); there are 4 regions, so there are 4 place files for each
geography-format-year.

#### Formats

* Monthly - "YYMM"
    * Current month - "c"
    * Year-to-date - "y"
    * Monthly cumulative - "r"
* Annual - "YYYY"
    * Annual summary - "a"
    * Annual revised - "r"

#### Regions

*For place files only.*

* Midwest - "mw"
* Northeast - "ne"
* South - "so"
* [West](https://media.giphy.com/media/f8HlgFSGiCpkk/giphy.gif) - "we"

The name of the BPS file tells you everything you need to know about it. 
They all follow the following template:

>\<geography or region>\<YYMM or YYYY>\<format>.txt

The state-level annual file for 2017 is called `st2017a.txt`. The county-level
current month file for January 2016 is called `co1601c.txt`. The MSA-level
year-to-date file for December 2000 is called `ma0012y.txt`. See the pattern?

Place files are special. The place-level monthly cumulative file for
February 1993 for the Northeast region is called `ne9302r.txt`. The place-level
annual summary file for 1990 for the West region is called `we1990a.txt`.

## Installation

You can install the released version of houser from [Github](https://www.github.com/everet/houser) with:

```{r, eval=FALSE}
library(devtools)
devtools::install_github("everetr/houser")
```

The "master" branch, installed by default above, is the latest stable release.
The "devel" branch is the latest development version, still being tested and 
therefore not recommended.

## Functions

* `bps_get()` - Download BPS data for specified geography(ies), format(s), year(s), and region(s) (if applicable).
* `bps_read()` - Load BPS data. Option to repair column names.
* `bps_doc()` - Open BPS documentation for a specified geography in your system's default PDF viewer.

## Usage

```{r, eval=FALSE}
library(houser)

# Download annual data for one geography, two years.
bps_get(path = ".", geography = "state", years = c(2010, 2017))

# Downloading place data requires a `region` argument.
bps_get(path = ".", geography = "place", region = "ne", years = 2016:2017)

# If the `return_log_df` argument is TRUE, bps_get will download the data in 
# the background, as usual, AND return a data frame containing metadata for all
# files that were downloaded.
bps_log = bps_get(path = ".", 
                  geography = "place", region = "ne", years = 2016:2017,
                  return_log_df = T)
```

## Contributor Code of Conduct

Please note that the 'houser' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
